---
title: "Data & Methods"
author: "by: Nina Branten"
bibliography: references.bib
---



```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide', eval=FALSE}
install.packages("remotes")
remotes::install_github("rlesur/klippy")
``` 

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
library(rgl)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

---


# Data and methods
&nbsp;

## Data
&nbsp;

### Staff members sociology
The current study focusses on the collaboration networks of staff members working at the  sociology department of the Radboud University and Utrecht University. Names of the staff members are webscraped from the websites of the university, using the RSelenium package for the javascript tables of the website of Utrecht University.

&nbsp;

### Collaboration network
Data on the publications and citations of the staff members were webscraped from Google Scholar, using the Scholar package in R.
For the RSiena analysis, collaborations between staff-members within the sociology departments themselves is operationalized using names of co-authors on the publications of the staff members within the period 2016-2022.
Furthermore, to decide the level of ethnic diversity of the collaborator egonets, I also gathered information on co-author relationships outside the department is also needed. This network is operationalized using the names of co-authors on the publications of staffmembers since 2019. Because we only use information on the publications since 2019, we assume that the diversity of scholar’s network is relatively stable over time.

&nbsp;

### Ethnic diversity
Ethnicity of the scholars is decided looking at their last name. Scholars are divided into two groups: (0) a non-Dutch last name, and (1) a Dutch last name. When in doubt about the origin of a last name, the CBG family name database was used, containing information on the geographical distribution and origin of last names.
The ethnic diversity of the egonet is operationalized as the percentage of ego’s co-authors since 2019 with non-Dutch last names.

&nbsp;

### Gender
Gender of the staff members is decided by looking at their first name, using the Meertens Voornamenbank. This a database contains information on the gender, popularity, geographical distribution and origin of all first names in The Netherlands.

&nbsp;

### Age
As data on the age of scholars was not easily available on the internet, we operationalized age as the period in which the scholar has been active in the academic world. Therefore age is seen as the year of the first publication of the staff member. 

&nbsp;

## Methods
First, several descriptive statistics of the networks are presented. A visual representation of the collaboration networks within the departments is made. Furthermore, several descriptive and frequency statistics are calculated regarding the ethnicity, age and gender of the ego’s, the ethnic diversity of the egonets and the degrees of the networks. Moreover, to look into the segregation of the networks, we compared intra- and intergroup densities of both gender and ethnicity.
Second, we try to explain differences in ethnic diversity of the collaboration egonets by characteristics of the egos themselves; ethnicity, gender and age. Therefore, two linear regression models were conducted.
Last, to find further explanations for segregation in network layers we need to examine how and why collaboration ties within the departments are formed and how this is influenced by covariate effects and/or structural network effects. This is done using SIENA (Simulation Investigation for Empirical Network Analysis) in R with the package RSiena. Our data contains information on collaboration ties between staff-members within the department of three time-frames (2016-2017, 2018-2019 and 2020-2022).
RSiena models tie changes as the results of actions by actors. It is assumed that each time only one actor is allowed to make changes, and this actor is only able to change one tie at a time. The actor who is allowed to make changes evaluates the networks structures on its attractiveness. This is determined by the evaluation function. The actor is most likely to make changes resulting in a network with the highest level of attractiveness. 
