---
title: "Research question attempt 1"
author: "by: Nina Branten"
bibliography: references.bib
---



```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide', eval=FALSE}
install.packages("remotes")
remotes::install_github("rlesur/klippy")
``` 

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
library(rgl)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

---

# Research question

*To what extent does international co-authorship between the West and developing countries follow neo-colonial patterns?*

---
---
**Relevant research papers:**

-	Boshoff, N. (2009). Neo-colonialism and research collaboration in Central Africa. Scientometrics, 81, 413-434. DOI: 10.1007/s11192-008-2211-8
-	Glänzel, W. & Schubert, A. (2005). Analysing scientific networks through co-authorship. In: H. F. Moed, W. Glänzel, & U. Schmoch (Eds.), Handbook of quantitative science and technology research (pp. 257 – 276). Dordrecht: Springer.
-	Nagtegaal, L. W. & Bruin, R. E. de. (1994). The French connection and other neo-colonial patterns in the global network of science. Research Evaluation, 4(2), 119-127. DOI: 10.1093/rev/4.2.119
