---
title: "Lab journal"
author: "by: Nina Branten"
bibliography: references.bib
---



```{r, install remotely, globalsettings, echo=FALSE, warning=FALSE, results='hide', eval=FALSE}
install.packages("remotes")
remotes::install_github("rlesur/klippy")
``` 

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
library(rgl)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```




----


# research questions  

Question 1: To what extent do the co-authorships  networks of the sociology department of the Radboud University and the sociology department of Utrecht University differ in their level of ethnic diversity?

Question 2: To what extent does the ethnic diversity of co-authorship networks of younger researchers differ from the ethnic diversity of co-authorship networks of older researchers?

Question 3: To what extent does the ethnic diversity of co-authorship networks of researchers with a migration background differ from the ethnic diversity of co-authorship networks of researchers without a migration background?

Question 4: To what extent are researchers influenced by each other in the formation of co-authorships with researchers from a different ethnic background?